A Distribuited Segment-based Approach for Topology-agnostic Nano Networks Routing

### 2013-04-15 PAPER/paper.tex
Fixed underscores

\author{
  \IEEEauthorblockN{Vincenzo Catania} \IEEEauthorblockA{Affiliation3\\ email3}} \and
  \IEEEauthorblockN{Davide Patti} \IEEEauthorblockA{Affiliation2\\ email2}

@@

link_visited[]: an array
of values representing information about attached links, that is, the
ID of the segment owning each link. When NULL, a link has not yet been
marked as visited. TDB: do we need to introduce some value for the
“bridge” status ? (e.g. a node connecting the terminal node of a
subnet with the starting one of another)

link_tvisited[]: an array of
values representing information about attached links, e.g. the ID of
the segment for which the link is candidate. When NULL, the link is
currently not tvisited.  starting: boolean, true if the node it's a
starting node for the subnet terminal:boolean, true if the node it's a
terminal node for the subnet.  subnet:the subnet the node belongs to.
If NULL, none.

@@

STARTING_SEGMENT_REQUEST packet should be injected in order to start
the whole process.

ACTIVE_SEARCHING: a node from which a find process of new segments has
been started and not yet cancelled or confirmed. This happens when all
the following conditions are matched: the node is marked as visited,
that is visited = TRUE

CHECK: segID is set to some value reflecting
the segment request sent there is a link that is being investigated by
the node, that is a tvisited_link marked with the same id of the node

CANDIDATE: a node currently candidate for belonging to some segment
with id segid, not being itself the node from which the find process
was started. The following conditions should be matched: tvisited =
true, with segID=X different from node's id.  Two links set as
link_tvisited with id X. Let’s refer them as  links 'i' and 'j'. Then
link_tvisited[i] was set = X when the find process from an adjacent
node reached the current node, while link_tvisited[j] has been set to
X when the node itself started to investigate its free links for the
segment request with id X. Note that while investigating its attached
links, if the find process fails along path 'j', the current node can
investigate another free node 'k' if suitable.

CANDIDATE_STARTING:
same as above, but the node is currently being considered as candidate
for starting segment. Two main differences: Can confirm
STARTING_SEGMENT_CONFIRM with the same segID when receiving a normal
SEGMENT_REQUEST: Overwrites its status setting it to CANDIDATE for the
segID frees the links previously set as tvisited during flooding

ASSIGNED: a node for which the segment has been determined.  The
segment segID attribute value is set to some id X different from NULL.
The visited value is true NOTE: an ASSIGNED state is a quick temporary
state, since it becomes ACTIVE_SEARCHING if a not (visited/tvisited)
link is suitable for searching new segments (see next_visited
procedure on paper)

@@

SEGMENT_REQUEST: used to search candidates for a segment (not the
first of the subnet, managed separately as a starting segment).

SEGMENT_CONFIRM: used to create a segment. When received, if the node
status is tvisited with a segID corresponding to the one indicated in
the packet, the node learns to belong to the segment associated to the
id. Further, it should forward this packet to the link where the
original SEGMENT_REQUEST packet came from, so that all candidate nodes
can learn the segment id they belong to. The following LED shoud be
updated: The state of the node changes from tvisited to visited.  the
link_tvisited previously set with the segID should be invalidated the
link_visited variable associated to the incoming confirm should be set
as segID. Note that if the request was flooded across different links,
the link_visited variable associated to the other paths shouldn’t be
set.  the link_visited variable associated to the link from which the
request was originated should be set segID

STARTING_SEGMENT_REQUEST: the request for the first segment of a
subnet should be differentiated since all nodes (except the starting
one) must forward the packet using a flooding mechanism. This is
necessary since the request packet must reach the starting node, which
is the only node marked as visited. As stated elsewhere, it should
remembered that the first segment of each subnet it's  a loop that
beginning/ending on the starting node.

SEGMENT_CANCEL: used when the process of searching a segment along a
certain path fails. When a node discovers that the path that includes
the node itself is no more suitable, a SEGMENT_CANCEL is sent back
across the link from which the SEGMENT_REQUEST was received.  Possible
reasons for that: The node it’s already candidate for another segment
The node it’s free but can’t find a free link (according the
cyclelinks timeout) The node it’s tvisited for the same request id,
previously forwarded that request along some free direction, but now
it’s receiving a a cancel request and has no more free links

@@

Injecting request: all nodes have status FREE, except for a
“bootstrap node” with status BOOTSTRAP, set by some signal from an
upper layer via. This bootstrap node sets itself as “visited” and
changes its status to ACTIVE_SEARCHING, sending a
STARTING_SEGMENT_REQUEST across one of its not (visited/tvisited)
links and.

Flooding: Each node receiving a STARTING_SEGMENT_REQUEST,
if not (visited/tvisited),  forwards  to the attached links that are
not already marked as visited/tvisited using a flooding mechanism.
Each of these links is then marked as tvisited with the segment id
associated to the packet. Note that a node that has already received a
STARTING_SEGMENT_REQUEST packet can simply ignore further packets
associated to the same request, since has already contributed to the
flooding. Note also that each of these packets has a max_segment_hops
field to prevent packets from travelling undefinitively.

setup of the starting-segment: when this STARTING_SEGMENT_REQUEST packet reaches
the starting node (from a different link, of course), the starting
segment is found. Then the starting node sends back a SEGMENT_CONFIRM
packet with the same id along the link from which it received the
STARTING_SEGMENT_REQUEST. Each node do same by changing its own status
from tvisited to visited. So the confirmation packet is sent back from
node to node and the starting segment is created.

Note: A node knows to be tvisited due a  STARTING_SEGMENT_REQUEST.
Then, when receiving a SEGMENT_REQUEST for other (not the starting)
segment, simply knows to cancel the tvisited status of the node and
its links, since if a segment arrived it means that a starting segment
has already been found a there’s no need to keep the tvisited status
for those broadcast status. What happens to nodes not receiving these
segment request? they simply remain tvisited and after a timeout reset
their state as free. But they could also remain tvisited for the
starting segment, since this doesn’t affect their behaviour for future
segment requests.

@@

Injecting request: A node in the ASSIGNED state can initiate a search for a
segment, by sending a SEGMENT_REQUEST with an across one of its free
links. Note that, since this is not the starting segment, in this case
the packet should not reach the initiator, but just another visited
node

setup of the segment: The find process is successfull when a visited
node receives the SEGMENT_REQUEST packet. Then, a SEGMENT_CONFIRM is
sent back along the path that originated the request.  Failing while
searching a segment: a node received a SEGMENT_REQUEST packet but
matched one of the following conditions: the node is free but has no
more suitable free links (can’t forward the SEGMENT_REQUEST) the node
is tvisited candidate for another find process the request packet
exceeded the max_segment_hops limit.  In all these cases the node
sends back a SEGMENT_CANCEL along the proper link modifying its state
to free. Each of the subsequent nodes will forward this SEGMENT_CANCEL
only if no more free links can be explored starting from them.

\subsection{Intra-node vs Inter-node parallelism}
In the processes described above, we assumed that each node in the
READY_SEARCHING state can start a find segment process by injecting a
SEGMENT_REQUEST. However, a more accurate decision when defining DiSR
should be made in order to decide whether or not the node must
actually perform this action. 
The point above is is strictly related to the general question: which
level of “parallelism” should we allow in DiSR ? The adopted approach
is that, although the nature of DiSR itself is intrinsicly parallel,
the use of parallelism makes things work in a more complex way. In
other words, DiSR is parallel when needed, but it does not exploit
parallelism as an “improving feature”. When not needed, things should
be serialized. For example: when a node is “ready searching” could
start several find segment process associated to the same segment, one
for each free link. But serializing this by investigating the free
links in order could be a simplest solution. We refer to this saying
that we avoid intra-node parallelism.  Note that while intra-node
parallelism can be avoided, trying to avoid inter-node parallelism
could be more complex than allowing the parallelism itself. Imagine
for example the effort of trying to coordinate nodes so that an unique
finding node process is actually running in the whole subnet. Thus, in
contrast to the intra-node parallelism, the inter-node parallelism is
a structural property of the DiSR algorithm and should not be avoided.
%------------------------------------------------------------------------------

@@

\subsection{Receiving a STARTING_SEGMENT_REQUEST}

When a node receives a STARTING_SEGMENT_REQUEST, with id = X,  there are the following cases:

The node is visited with the same segID = x, that is it was the
initiator of the request (set visited by definition): a first segment
was found and should be confirmed sending back a SEGMENT_CONFIRM
packet.

@@

The node it's visited, with different id in the packet.
CHECK:  it is not the initiator of the starting segment request, but
is already assigned to some segment!? assuming again we are discussing
of nodes belonging to the same subnet, this should NOT happen, since
the only visited node of the subnet should be the initial node from
which the STARTING_SEGMENT_REQUEST originated. Note again that we are
assuming that if the the node is visited or tvisited with the id of
another subnet can simply ignore the request.  The node it's tvisited
but it's not the initiator of the request (different id): same as
above, if we are assuming nodes of the same subnet, this shouldn’t
happen. If the node is of another subnet, can simply ignore the event.

\subsection{Receiving a SEGMENT_REQUEST}
When a node receives a SEGMENT_REQUEST, the are the following cases:

The node is visited (and belongs to the same subnet): a segment was
found and it should be confirmed sending back a SEGMENT_CONFIRM
packet.

The node is tvisited: it should discard the packet sending
back a SEGMENT_CANCEL

The node it's not visited/tvisited and has free
links: it marks itself as tvisited and forwards the SEGMENT_REQUEST
(one request at time) to each of its free links, according to some
ordering

CHECK: the difference between confirming a STARTING_SEGMENT_REQUEST
and confirming a SEGMENT_REQUEST is that in the first case the node
itself it's included in the segment , while in the second case the
receiving node DOES NOT belong to the segment found. The motivation is
that in the second case the node was already marked as visited because
of being assigned to some segment found before, while in the
starting-segment case, by definition, the first segment of a subnet
begins with the starting node, which is marked as visited by the SR
algorithm when searching for the first (starting) segment. Remember
that the first segment of a subnet is a loop beginning with the
starting node.

\subsection{Receiving a SEGMENT_CANCEL}
When a node receives a SEGMENT_CANCEL packet it means that searching
for a segment along that path was unsuccessful. But if the node still
has other free links to try, it should forward a SEGMENT_REQUEST to
the next link (next according to some internal ordering). So a node
forwards back the SEGMENT_CANCEL packet along the link that originated
the SEGMENT_REQUEST packet ONLY when there are no more free links to
try. If this is the case, the node modifies its status from tvisited
to free and fordwards the SEGMENT_CANCEL packet to the link from which
the request was received. The process stops when the SEGMENT_CANCEL
packet reach the starting node that originated the request.

\section{Nanoxim and simulation environment}
% - Architecture of the nanoxim simulator
% - how simulations were conducted
...

>>>>

\author{
  \IEEEauthorblockN{a} \IEEEauthorblockA{Affiliation3\\ email3} \and
  \IEEEauthorblockN{b} \IEEEauthorblockA{Affiliation2\\ email2} }

@@

link\_visited[]: an array
of values representing information about attached links, that is, the
ID of the segment owning each link. When NULL, a link has not yet been
marked as visited. TDB: do we need to introduce some value for the
“bridge” status ? (e.g. a node connecting the terminal node of a
subnet with the starting one of another)

link\_tvisited[]: an array of
values representing information about attached links, e.g. the ID of
the segment for which the link is candidate. When NULL, the link is
currently not tvisited.  starting: boolean, true if the node it's a
starting node for the subnet terminal:boolean, true if the node it's a
terminal node for the subnet.  subnet:the subnet the node belongs to.
If NULL, none.

@@

STARTING\_SEGMENT\_REQUEST packet should be injected in order to start
the whole process.

ACTIVE\_SEARCHING: a node from which a find process of new segments has
been started and not yet cancelled or confirmed. This happens when all
the following conditions are matched: the node is marked as visited,
that is visited = TRUE

CHECK: segID is set to some value reflecting
the segment request sent there is a link that is being investigated by
the node, that is a tvisited\_link marked with the same id of the node

CANDIDATE: a node currently candidate for belonging to some segment
with id segid, not being itself the node from which the find process
was started. The following conditions should be matched: tvisited =
true, with segID=X different from node's id.  Two links set as
link\_tvisited with id X. Let’s refer them as  links 'i' and 'j'. Then
link\_tvisited[i] was set = X when the find process from an adjacent
node reached the current node, while link\_tvisited[j] has been set to
X when the node itself started to investigate its free links for the
segment request with id X. Note that while investigating its attached
links, if the find process fails along path 'j', the current node can
investigate another free node 'k' if suitable.

CANDIDATE\_STARTING:
same as above, but the node is currently being considered as candidate
for starting segment. Two main differences: Can confirm
STARTING\_SEGMENT\_CONFIRM with the same segID when receiving a normal
SEGMENT\_REQUEST: Overwrites its status setting it to CANDIDATE for the
segID frees the links previously set as tvisited during flooding

ASSIGNED: a node for which the segment has been determined.  The
segment segID attribute value is set to some id X different from NULL.
The visited value is true NOTE: an ASSIGNED state is a quick temporary
state, since it becomes ACTIVE\_SEARCHING if a not (visited/tvisited)
link is suitable for searching new segments (see next\_visited
procedure on paper)

@@

SEGMENT\_REQUEST: used to search candidates for a segment (not the
first of the subnet, managed separately as a starting segment).

SEGMENT\_CONFIRM: used to create a segment. When received, if the node
status is tvisited with a segID corresponding to the one indicated in
the packet, the node learns to belong to the segment associated to the
id. Further, it should forward this packet to the link where the
original SEGMENT\_REQUEST packet came from, so that all candidate nodes
can learn the segment id they belong to. The following LED shoud be
updated: The state of the node changes from tvisited to visited.  the
link\_tvisited previously set with the segID should be invalidated the
link\_visited variable associated to the incoming confirm should be set
as segID. Note that if the request was flooded across different links,
the link\_visited variable associated to the other paths shouldn’t be
set.  the link\_visited variable associated to the link from which the
request was originated should be set segID

STARTING\_SEGMENT\_REQUEST: the request for the first segment of a
subnet should be differentiated since all nodes (except the starting
one) must forward the packet using a flooding mechanism. This is
necessary since the request packet must reach the starting node, which
is the only node marked as visited. As stated elsewhere, it should
remembered that the first segment of each subnet it's  a loop that
beginning/ending on the starting node.

SEGMENT\_CANCEL: used when the process of searching a segment along a
certain path fails. When a node discovers that the path that includes
the node itself is no more suitable, a SEGMENT\_CANCEL is sent back
across the link from which the SEGMENT\_REQUEST was received.  Possible
reasons for that: The node it’s already candidate for another segment
The node it’s free but can’t find a free link (according the
cyclelinks timeout) The node it’s tvisited for the same request id,
previously forwarded that request along some free direction, but now
it’s receiving a a cancel request and has no more free links

@@

Injecting request: all nodes have status FREE, except for a
“bootstrap node” with status BOOTSTRAP, set by some signal from an
upper layer via. This bootstrap node sets itself as “visited” and
changes its status to ACTIVE\_SEARCHING, sending a
STARTING\_SEGMENT\_REQUEST across one of its not (visited/tvisited)
links and.

Flooding: Each node receiving a STARTING\_SEGMENT\_REQUEST,
if not (visited/tvisited),  forwards  to the attached links that are
not already marked as visited/tvisited using a flooding mechanism.
Each of these links is then marked as tvisited with the segment id
associated to the packet. Note that a node that has already received a
STARTING\_SEGMENT\_REQUEST packet can simply ignore further packets
associated to the same request, since has already contributed to the
flooding. Note also that each of these packets has a max\_segment\_hops
field to prevent packets from travelling undefinitively.

setup of the starting-segment: when this STARTING\_SEGMENT\_REQUEST packet reaches
the starting node (from a different link, of course), the starting
segment is found. Then the starting node sends back a SEGMENT\_CONFIRM
packet with the same id along the link from which it received the
STARTING\_SEGMENT\_REQUEST. Each node do same by changing its own status
from tvisited to visited. So the confirmation packet is sent back from
node to node and the starting segment is created.

Note: A node knows to be tvisited due a  STARTING\_SEGMENT\_REQUEST.
Then, when receiving a SEGMENT\_REQUEST for other (not the starting)
segment, simply knows to cancel the tvisited status of the node and
its links, since if a segment arrived it means that a starting segment
has already been found a there’s no need to keep the tvisited status
for those broadcast status. What happens to nodes not receiving these
segment request? they simply remain tvisited and after a timeout reset
their state as free. But they could also remain tvisited for the
starting segment, since this doesn’t affect their behaviour for future
segment requests.

@@

Injecting request: A node in the ASSIGNED state can initiate a search for a
segment, by sending a SEGMENT\_REQUEST with an across one of its free
links. Note that, since this is not the starting segment, in this case
the packet should not reach the initiator, but just another visited
node

setup of the segment: The find process is successfull when a visited
node receives the SEGMENT\_REQUEST packet. Then, a SEGMENT\_CONFIRM is
sent back along the path that originated the request.  Failing while
searching a segment: a node received a SEGMENT\_REQUEST packet but
matched one of the following conditions: the node is free but has no
more suitable free links (can’t forward the SEGMENT\_REQUEST) the node
is tvisited candidate for another find process the request packet
exceeded the max\_segment\_hops limit.  In all these cases the node
sends back a SEGMENT\_CANCEL along the proper link modifying its state
to free. Each of the subsequent nodes will forward this SEGMENT\_CANCEL
only if no more free links can be explored starting from them.

\subsection{Intra-node vs Inter-node parallelism}
In the processes described above, we assumed that each node in the
READY\_SEARCHING state can start a find segment process by injecting a
SEGMENT\_REQUEST. However, a more accurate decision when defining DiSR
should be made in order to decide whether or not the node must
actually perform this action. 
The point above is is strictly related to the general question: which
level of “parallelism” should we allow in DiSR ? The adopted approach
is that, although the nature of DiSR itself is intrinsicly parallel,
the use of parallelism makes things work in a more complex way. In
other words, DiSR is parallel when needed, but it does not exploit
parallelism as an “improving feature”. When not needed, things should
be serialized. For example: when a node is “ready searching” could
start several find segment process associated to the same segment, one
for each free link. But serializing this by investigating the free
links in order could be a simplest solution. We refer to this saying
that we avoid intra-node parallelism.  Note that while intra-node
parallelism can be avoided, trying to avoid inter-node parallelism
could be more complex than allowing the parallelism itself. Imagine
for example the effort of trying to coordinate nodes so that an unique
finding node process is actually running in the whole subnet. Thus, in
contrast to the intra-node parallelism, the inter-node parallelism is
a structural property of the DiSR algorithm and should not be avoided.
%------------------------------------------------------------------------------

@@

\subsection{Receiving a STARTING\_SEGMENT\_REQUEST}

When a node receives a STARTING\_SEGMENT\_REQUEST, with id = X,  there are the following cases:

The node is visited with the same segID = x, that is it was the
initiator of the request (set visited by definition): a first segment
was found and should be confirmed sending back a SEGMENT\_CONFIRM
packet.

@@

The node it's visited, with different id in the packet.
CHECK:  it is not the initiator of the starting segment request, but
is already assigned to some segment!? assuming again we are discussing
of nodes belonging to the same subnet, this should NOT happen, since
the only visited node of the subnet should be the initial node from
which the STARTING\_SEGMENT\_REQUEST originated. Note again that we are
assuming that if the the node is visited or tvisited with the id of
another subnet can simply ignore the request.  The node it's tvisited
but it's not the initiator of the request (different id): same as
above, if we are assuming nodes of the same subnet, this shouldn’t
happen. If the node is of another subnet, can simply ignore the event.

\subsection{Receiving a SEGMENT\_REQUEST}
When a node receives a SEGMENT\_REQUEST, the are the following cases:

The node is visited (and belongs to the same subnet): a segment was
found and it should be confirmed sending back a SEGMENT\_CONFIRM
packet.

The node is tvisited: it should discard the packet sending
back a SEGMENT\_CANCEL

The node it's not visited/tvisited and has free
links: it marks itself as tvisited and forwards the SEGMENT\_REQUEST
(one request at time) to each of its free links, according to some
ordering

CHECK: the difference between confirming a STARTING\_SEGMENT\_REQUEST
and confirming a SEGMENT\_REQUEST is that in the first case the node
itself it's included in the segment , while in the second case the
receiving node DOES NOT belong to the segment found. The motivation is
that in the second case the node was already marked as visited because
of being assigned to some segment found before, while in the
starting-segment case, by definition, the first segment of a subnet
begins with the starting node, which is marked as visited by the SR
algorithm when searching for the first (starting) segment. Remember
that the first segment of a subnet is a loop beginning with the
starting node.

\subsection{Receiving a SEGMENT\_CANCEL}
When a node receives a SEGMENT\_CANCEL packet it means that searching
for a segment along that path was unsuccessful. But if the node still
has other free links to try, it should forward a SEGMENT\_REQUEST to
the next link (next according to some internal ordering). So a node
forwards back the SEGMENT\_CANCEL packet along the link that originated
the SEGMENT\_REQUEST packet ONLY when there are no more free links to
try. If this is the case, the node modifies its status from tvisited
to free and fordwards the SEGMENT\_CANCEL packet to the link from which
the request was received. The process stops when the SEGMENT\_CANCEL
packet reach the starting node that originated the request.

\section{Nanoxim and simulation environment}

The simulator Nanoxim is a SystemC tool based on a minimal version of
the Noxim Network-on-Chip simulator. It's not a "faster noxim", but a
drastically reduced noxim the will be able to simulate a random
network on nodes. Current differences from noxim:

removed all routing algorithms that depend on a topology (only XY left
to test meshes) removed flit management, nodes only exchange packets
(e.g. no wormhole)

Future features:

topology agnostic able to test different nano-oriented communication
mechanism
